Generality in food webs scales with species richness, not latitude

### 2014-07-29 manuscript/latitudepaper.tex
Daniel corrections.

\subsection*{Data Set}
We compiled a list of 347 empirical food webs from an online database (GlobalWeb webs \citep{GlobalWeb}), supplemented 
with 31 food webs used to study the relationship between consumer and resource body sizes (SizeWebs \citep{Brose2006})
and seven food webs used to study the impact of parasites on network structure (ParWebs \citep{Dunne2013}).
Of the GlobalWeb webs, 122 were rejected because their original source could not be found or was unpublished (53), they were source or sink
webs rather than descriptions of an entire community (13), were focused on plant-pollinator, host-parasitoid, or 
competitive interactions (8+34+2=44), described inferred interactions in an extinct community (2), were ``generalized 
schemes'' rather than being based on empirical observation (8), or because it was not clear which of a variety of 
described sites the published food web represented (2). (See Table S1 for a list of rejected webs and reasons for their
exclusion.) For three GlobalWeb webs which included parasites as a minor component of the community, as well as the 
seven ParWebs webs, we included a modified version of each web which included only free-living species and the 
interactions between them \citep{Dunne2013}. For each of the 263 webs included in the final database, we recorded the
study site latitude if possible or, where no latitude was given, obtained estimated coordinates of the described
study site using Google Earth \citep{GoogleEarth}. Webs were also classified according to ecosystem type (terrestrial,
lake, stream, estuary, or marine), year of publication (to control for changing standards of data collection over time \citep{Dunne2006}), and whether or not humans were included (11 webs included humans).

@@

\begin{eqnarray}
S_{i} = \alpha_{0} + \alpha_{1} \Delta_{i} + \epsilon_{i} 
\end{eqnarray}

where $S_{i}$ is the species richness of web $i$, $\Delta_{i}$ its absolute latitude (degrees north or south 
regardless of direction), and $\epsilon_{i}$ a residual error term.
These simple linear models, fit using the function lm in R \citep{R} package stats, were used to confirm that there are indeed latitudinal gradients in food-web structure in this data set.

@@

However, the two forms of the relationship imply different error structures: normally-distributed, additive error in 
\ref{Power} and lognormal, multiplicative error in \ref{Loglog}. As there is no intuitive reason to believe that our
data have one error distribution over another, we follow the recommendations in \citet{Xiao2011} and compared the two
model formulations explicitly. The model with the error distribution most resembling that observed in the empirical
data was then used to test for potential effects of latitude.

@@

@@

Once the best formulation of the relationship between each property and $S$ had been determined, we then assessed
the impact of latitude on this relationship. There are two ways that properties might be affected by latitude:
1) the slope of the scaling relationship with $S$ might vary over latitude or 2) the residuals $\epsilon$ might
vary over latitude. To account for both possibilities we included terms for the effect of latitude on the scaling
exponent $\beta$ in equation \ref{Power} as well as on the residual. In addition, as other studies have suggested that 
food webs in different ecosystem types have characteristic structural differences, we included a dummy variable for 
each ecosystem type (stream, lake, terrestrial, or marine, with estuary used as the `intercept' ecosystem type). 
We also considered other trends in the data that could affect relationships between latitude and species richness.
To account for changing standards in data collection over time \citep{Dunne2006} we included an effect of year of
publication. As 11 of our XX webs included humans, and therefore might have different properties than other networks,
we included a factorial effect of the presence of humans. Finally, some of our webs describe the same area in different
seasons or different years. To account for likely correlation among these webs, we included a random effect of site ID.

@@

\begin{equation}
\label{PowerLat}
Z_{i}=\alpha S^{\beta_{0}+\beta_{1}\Delta_{i}+\beta_{3}\Sigma_{i} + \beta_{4}\Lambda_{i} + \beta_{5}T_{i} + \beta_{6}M_{i} 
+ \beta_{7}\Delta_{i}\Sigma_{i} + \beta_{8}\Delta_{i}\Lambda_{i} + \beta_{9}\Delta_{i}T_{i} + \beta_{10}\Delta_{i}M_{i}}
+ \rho{1}\Delta_{i} + \rho_{2}\Upsilon_{i} + \rho_{3}H_{i} + P_{i} + \epsilon_{i} 
\end{equation}

where $\delta$ is the absolute latitude of food web $i$ (degrees from the equator, regardless of direction),
$\Sigma_{i}$, $\Lambda_{i}$, $T_{i}$, and $M_{i}$ are 
factors indicating whether web $i$ is a stream, lake, 
terrestrial, or marine food web respectively (e.g., $\Sigma_{i}$=1 for 
a stream food web, 0 otherwise). If $\Sigma_{i}$, $\Lambda_{i}$, $T_{i}$, and $M_{i}$ are all 0, web $i$ 
describes an estuary food web. $\Upsilon_{i}$ is the year web $i$ was published, $H_{i}=1$ if web $i$ contains humans, 0 otherwise, and $P_{i}$ is a random effect of 
site ID. Note that all fixed effects $\beta$ refer to factors acting on the scaling exponent between $Z$ and $S$
while fixed effects $\rho$ refer to factors acting on the residual error.

@@

\begin{equation}
\label{LogLat}
Z_{i}= log(\alpha)+\beta_{0}log(S) + \beta_{1}\Delta_{i}log(S) +\beta_{3}\Sigma_{i}log(S) + \beta_{4}\Lambda_{i}log(S)  + \beta_{5}T_{i}log(S)  + \beta_{6}M_{i}log(S)  
+ \beta_{7}\Delta_{i}\Sigma_{i}log(S)  + \beta_{8}\Delta_{i}\Lambda_{i}log(S)  + \beta_{9}\Delta_{i}T_{i}log(S)  + \beta_{10}\Delta_{i}M_{i}log(S) 
+ \rho_{1}\Delta_{i} + \rho_{2}\Upsilon_{i}) + \rho_{3}H_{i} +\epsilon_{i}
\end{equation}

As in the power-law formulation, fixed effects $\beta$ 
affect the scaling exponent and fixed effects $\rho$ 
affect the (lognormal) residual.

We calculated the AIC of the chosen maximal model, as well as the AIC's of a suite of candidate simplified models.
Simplified models were obtained by systematically removing all possible combinations of terms from the full model.
The term for species richness was not removed from a model if any of the terms affecting the exponent $\beta$ were 
included in the model. The model with the lowest AIC was selected as the best-fitting model. For power-law formulations
this simplification was performed by hand. For logarithmic formulations, this simplification was performed using the
R \citep{R} function dredge from package MuMIn \citep{MuMIn}. We then used the R \citep{R} function lmer
from the package lmerTest \citep{lmerTest} to estimate the standardized effects ($\beta$s and $\rho$s) for each fixed effect in the 
best-fitting models, as well as their corresponding $p$-values.

@@

\multirow{2}{*}Property & \multicolumn{2}{|c}{Intercept} & \multicolumn{2}{|c}{Slope} \\
                      & $\alpha$ & $p$-value & $\alpha$ & $p$-value \\
\hline
$S$ & 23.10 & \textless0.001 & 0.284 & 0.040 \\
$Z$ & 2.566 & \textless0.001 & 0.020 & 0.180 \\
$G$ & 3.269 & \textless0.001 & 0.049 & 0.014 \\
$V$ & 2.566 & \textless0.001 & 0.020 & 0.180 \\
\end{tabular}
\end{table}

@@

For models of $Z$, $G$, and $V$, the assumption of multiplicative log-normal error was better supported (Table \ref{Errorfits}). We therefore use only models with logarithmic form for the remainder of the paper.

\begin{table}[!t]

\caption{AIC's for model \ref{Power}, assuming normal error, and model \ref{Loglog}, assuming lognormal error.}
\label{Errorfits}
\begin{tabular}{l | l l | c}
Property & Normal error & Lognormal error & Model selected \\
\hline
$Z$ & 1151 &  767 & \ref{Loglog} \\
$G$ & 1266 & 1037 & \ref{Loglog} \\
$V$ & 1151 & 767 & \ref{Loglog} \\
\end{tabular}
\end{table}

@@

\caption{Standard effects and their $p$-values for the best-fitting models of $Z$, $G$, and $V$.}
\label{Bestfits}
\begin{tabular}{l | l l | l l | l l | l l }
\multirow{2}{*}{Property} & \multicolumn{2}{c}{Intercept} & \multicolumn{2}{c}{log(Species)} & \multicolumn{2}{c}{log(Species)$\Lambda$} & \multicolumn{2}{c}{$\Upsilon$} \\
& $\beta$ & $p$-value &  $\beta$ & $p$-value &  $\beta$ & $p$-value &  $\beta$ & $p$-value  \\
\hline
$Z$ & -0.516 & \textless0.001 & 0.662 & \textless0.001 & 0.099 & \textless0.001 & \multicolumn{2}{c}{NA} \\
$G$ & -9.999 & \textless0.001 & 0.547 & \textless0.001 & \multicolumn{2}{c}{NA} & 0.005 & \textless0.001 \\
$V$ & -0.516 & \textless0.001 & 0.662 & \textless0.001 & 0.099 & \textless0.001 & \multicolumn{2}{c}{NA} \\
\end{tabular}
\end{table}

@@

As the latitudinal gradient in species richness in this data set ran in the
opposite direction to the general trend for \citep{}, we conclude that it is
likely an artifact of the data set. Academic institutions have the longest
history in temperate latitudes, and it may be that communities at these
latitudes are better understood or more completely sampled. As such, it does not seem likely that the increase in
generality with latitude that we observed is due to an underlying tendency for
species to pursue a wider variety of prey at higher latitudes. Rather, the
trend in generality can be more parsimoniously explained by the trend in
species richness.

@@

It is not our intention here to question the validity of the wide range of
studies exploring variation in network structure across changing environments.
However, if those studies have not taken into account changes in species
richness over environmental gradients, it is possible that observed changes in
food-web structure are due only indirectly to the environmental variables of
interest. Effects of environmental variables on network structure should
therefore be interpreted with caution unless effects of changing species
richness are made explicit.

>>>>

\subsection*{Data Set}

We compiled a list of 263 empirical food webs from
multiple sources (see Supplemental Information for web origins and selection
criteria). We recorded study site latitude from the original source where
possible or, where study sites were described but exact coordinates were not
given, obtained estimated coordinates using Google Earth \citep{GoogleEarth}.
Webs were also classified according to ecosystem type (terrestrial, lake,
stream, estuary, or marine), year of publication (to control for potential change
in standards of data collection over time \citep{Dunne2006}), and whether or not
humans were included.

% [[list in supplemental (347 GlobalWeb webs \citep{GlobalWeb}), supplemented 
% with 31 food webs used to study the relationship between consumer and resource body sizes (SizeWebs \citep{Brose2006})
% and seven food webs used to study the impact of parasites on network structure (ParWebs \citep{Dunne2013}).
% Of the GlobalWeb webs, 122 were rejected because their original source could not be found or was unpublished (53), they were source or sink
% webs rather than descriptions of an entire community (13), were focused on plant-pollinator, host-parasitoid, or 
% competitive interactions (8, 34, and 2, respectively), described inferred interactions in an extinct community (2), were ``generalized 
% schemes'' rather than being based on empirical observation (8), or because it was not clear which of a variety of 
% described sites the published food web represented (2). (See Table S1 for a list of rejected webs and reasons for their
% exclusion.) For three GlobalWeb webs which included parasites as a minor component of the community, as well as the 
% seven ParWebs webs, we included a modified version of each web which included only free-living species and the 
% interactions between them \citep{Dunne2013}.

@@

\begin{eqnarray}
S_{i} = \alpha_{0} + \alpha_{1} L_{i} + \epsilon_{i} 
\end{eqnarray}

where $S_{i}$ is the species richness of web $i$, $\Lambda_{i}$ its absolute latitude (degrees north or south 
regardless of direction), and $\epsilon_{i}$ a residual error term.
These simple linear models, fit using the function lm in R \citep{R} package stats, were used to confirm that there are indeed latitudinal gradients in food-web structure in this data set.

@@

@@

%%% TO SI

% Note that the two forms of the power-law relationship imply different error structures: normally-distributed, additive error in 
% \ref{Power} and lognormal, multiplicative error in \ref{Loglog}. As there is no intuitive reason to believe that our
% data have one error distribution over another, we follow the recommendations in \citet{Xiao2011} and compared the two
% model formulations explicitly. The model with the error distribution most resembling that observed in the empirical
% data was then used to test for potential effects of latitude.

@@

We then assessed
the impact of latitude on this relationship; specifically, the effect of latitude on the scaling exponent $\beta$. 
In addition to latitude, we included a factorial variable for ecosystem type (stream, lake, terrestrial, marine, or estuary) to account for recognized variation in food web structure across ecosystem types \citep{}.
% We also considered other trends in the data that could affect relationships between latitude and species richness.
% To account for changing standards in data collection over time \citep{Dunne2006} we included an effect of year of
% publication. As 11 of our XX webs included humans, and therefore might have different properties than other networks,
% we included a factorial effect of the presence of humans. 
Finally, some of our webs describe the same area in different
seasons or different years. To account for likely correlation among these webs, we included a random effect of site ID.

@@

\begin{equation}
\label{PowerLat}
Z_{i}=\alpha S_{i}^{\beta_{0}+\beta_{1}L_{i}+\beta_{3}E_{i}} + P_{i} + \epsilon_{i} 
\end{equation}

where $L_{i}$ is the absolute latitude of food web $i$ (degrees from the equator, regardless of direction),
$E_{i}$ indicates the ecosystem type of food web $i$, and $P_{i}$ is a random effect of 
site ID.

@@

\begin{equation}
\label{LogLat}
log(Z_{i})= log(\alpha)+\beta_{0}log(S_{i}) + \beta_{1}L*log(S_{i}) +\beta_{3}E*log(S_{i})+ P_{i} +\epsilon_{i}
\end{equation}

We fit equation \ref{LogLat}, as well as similar equations for $G$ and $V$ using R \citep{R} function lmer from package lme4
\citep{lmerTest}.
We calculated the AIC of the chosen maximal model, as well as the AIC's of a suite of candidate simplified models.
Simplified models were obtained by systematically removing all possible combinations of terms from the full model.
The term for species richness was not removed from a model if any of the terms affecting the exponent $\beta$ were 
included in the model. The model with the lowest AIC was selected as the best-fitting model. 
This simplification was performed using the
R \citep{R} function dredge from package MuMIn \citep{MuMIn}. We then used the R \citep{R} function lmer
from the package lmerTest \citep{lmerTest} to estimate the standardized effects ($\beta$s ) for each fixed effect in the 
best-fitting models, as well as their corresponding $p$-values.

@@

\multirow{2}{*}{Property} & \multicolumn{2}{|c}{Intercept} & \multicolumn{2}{|c}{Slope} \\
                      & $\alpha_{0}$ & $p$-value & $\alpha_{1}$ & $p$-value \\
\hline
$S$ & 23.10 & \textless0.001 & 0.284 & 0.040 \\
$Z$ & 2.566 & \textless0.001 & 0.020 & 0.180 \\
$G$ & 3.269 & \textless0.001 & 0.049 & 0.014 \\
$V$ & 2.566 & \textless0.001 & 0.020 & 0.180 \\
\end{tabular}
\end{table}

@@

% For models of $Z$, $G$, and $V$, the assumption of multiplicative log-normal error was better supported (Table \ref{Errorfits}). We therefore use only models with logarithmic form for the remainder of the paper.

% \begin{table}[!t]

% \caption{AIC's for model \ref{Power}, assuming normal error, and model \ref{Loglog}, assuming lognormal error.}
% \label{Errorfits}
% \begin{tabular}{l | l l | c}
% Property & Normal error & Lognormal error & Model selected \\
% \hline
% $Z$ & 1151 &  767 & \ref{Loglog} \\
% $G$ & 1266 & 1037 & \ref{Loglog} \\
% $V$ & 1151 & 767 & \ref{Loglog} \\
% \end{tabular}
% \end{table}

@@

\caption{Standard effects and their $p$-values for the best-fitting models of $Z$, $G$, and $V$.}
\label{Bestfits}
\begin{tabular}{l | l l | l l | l l | l l }
\multirow{2}{*}{Property} & \multicolumn{2}{c}{Intercept} & \multicolumn{2}{c}{log(Species)} & \multicolumn{2}{c}{log(Species)$\Lambda$} & \multicolumn{2}{c}{$\Upsilon$} \\
& $\beta$ & $p$-value &  $\beta$ & $p$-value &  $\beta$ & $p$-value &  $\beta$ & $p$-value  \\
\hline
$Z$ & -0.516 & \textless0.001 & 0.662 & \textless0.001 & 0.099 & \textless0.001 & \multicolumn{2}{c}{NA} \\
$G$ & -9.999 & \textless0.001 & 0.547 & \textless0.001 & \multicolumn{2}{c|}{NA} & 0.005 & \textless0.001 \\
$V$ & -0.516 & \textless0.001 & 0.662 & \textless0.001 & 0.099 & \textless0.001 & \multicolumn{2}{c}{NA} \\
\end{tabular}
\end{table}

@@

The latitudinal gradient in species richness in this data set ran in the opposite 
direction to the general trends for (list types of taxa). This may be due to the 
fact that latitudinal gradient in species richness are usually studied within a
taxonomic or trophic group (e.g., birds, vascular plants), whereas food webs 
measure many taxa concurrently. It is possible that while many taxa generally increase
towards the tropics, the number of species across all taxa that co-occur within an 
study site decreases. Alternatively, it is possible that by using trophic-species versions
of our food webs, we have artificially reversed the species-richness gradient. [[re-run analyses
with species-species webs]]. This would imply that, while there may be more species in the tropics,
many of those species tend to have identical predator and prey sets to co-occuring species. In either
case, it does not seem likely that the increase in
generality with latitude that we observed is due to an underlying tendency for
species to pursue a wider variety of prey at higher latitudes. Rather, the
trend in generality can be more parsimoniously explained by the trend in
species richness.

@@

It was not our initial intention to question the validity of the wide range of
studies exploring variation in network structure across changing environments \citep{}.
However, if those studies have not taken into account changes in species
richness over environmental gradients, it is possible that observed changes in
food-web structure are due only indirectly to the environmental variables of
interest. Effects of environmental variables on network structure should
therefore be interpreted with caution unless effects of changing species
richness are made explicit.
